Interisland movements and oceanic swimming of woodland caribou in Newfoundland

### 2020-02-24 paper/manuscript.Rmd
Update manuscript.Rmd

Caribou (*Rangifer tarandus*) are exceptional swimmers. Ample evidence exists
that caribou swim in streams, rivers, and lakes during migration [@Leblond2016],
to avoid predators [@Bergerud1985], and access islands during calving
[@Bergerud1990]. Even for adept swimmers like caribou, the energetic
expenditure of swimming for quadrupedal mammals is significantly higher than
walking or running <!--[@Fish1993]-->, while drowning is also possible [@Miller1985].
Despite a large number of coastal and island caribou herds, only a handful of
examples have documented caribou swimming in the ocean [e.g. @Miller1995; @Miller2002; @Jeffery2007; @Ricca2012].

The Fogo Island archipelago, off the coast of Newfoundland, Canada, is home to
approximately 300 woodland caribou. During routine fieldwork on May 30, 2017, we
observed an unmarked adult male caribou swim between Western and Eastern Indian
Islands (Figure 1), a swim of at least 470 m which took ~9 minutes (~52 m per
minute). This observation prompted us to investigate whether swimming behaviour
was common among GPS radio-collared caribou in the population (n = 29 adult
females; 1240g, GPS 4400M; Lotek Wireless Inc., Newmarket, Ontario, Canada). We
identified swimming events as two consecutive GPS (2-hour relocation rates)
locations from an individual occurred on different islands. Northeastern
Newfoundland typically experiences pack ice during winter and caribou are known
locally to travel between islands by walking over the ice. We restricted our GPS
data to the ice-free period of the year (April 1 - November 25) <!--confirm
dates --> and we identified three adult female caribou that collectively had 86
swimming events (Figure 2). In addition to our own visual and remotely sensed
observations, residents of Fogo Island have also reported observing caribou
swimming between islands on numerous occasions.

>>>>

Caribou (*Rangifer tarandus*) are exceptional swimmers. Ample evidence exists
that caribou swim in streams, rivers, and lakes during migration [@Leblond2016],
to avoid predators [@Bergerud1985], and access islands during calving
[@Bergerud1990]. Even for adept swimmers like caribou, the energetic
expenditure of swimming for quadrupedal mammals is significantly higher than
walking or running, <!--[@Fish1993]--> while drowning is also possible [@Miller1985].
Despite a large number of coastal and island caribou herds, only a handful of
examples have documented caribou swimming in the ocean [e.g. @Miller1995; @Miller2002; @Jeffery2007; @Ricca2012].

The Fogo Island archipelago, off the coast of Newfoundland, Canada, is home to
approximately 300 woodland caribou. During routine fieldwork on May 30, 2017, we
observed an unmarked adult male caribou swim between Western and Eastern Indian
Islands (Figure 1), a swim of at least 470 m which took ~9 minutes (~52 m per
minute). This observation prompted us to investigate whether swimming behaviour
was common among GPS radio-collared caribou in the population (n = 29 adult
females; 1240g, GPS 4400M; Lotek Wireless Inc., Newmarket, Ontario, Canada). We
identified swimming events as two consecutive GPS (2-hour relocation rates)
locations from an individual occurred on different islands. Northeastern
Newfoundland typically experiences pack ice during winter and caribou are known
locally to travel between islands by walking over the ice. We restricted our GPS
data to the ice-free period of the year (April 1 - November 25)<!--confirm dates
--> and we identified three adult female caribou that collectively had 86
swimming events over three years (Figure 2). In addition to our own visual and
remotely sensed observations, residents of Fogo Island have also reported
observing caribou swimming between islands on numerous occasions.
